---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(camtrapr)
```

`camtrapr` is designed to process photos from camera trapping projects into a form useful for further analysis in R. The functions are inspired by the Windows programs developed for the same purpose by the [Small Wild Cat Conservation Foundation](http://www.smallcats.org/). So, users of these tools will easily be able to transition to `camtrapr`.

The motivation behind creating this package was a desire to have an open source, user friendly, cross platform tool (i.e. not Windows only) that allows for a seamless R camera trapping workflow.

# Installation

```{r, eval=F}
install.packages("devtools")
devtools::install_github("mstrimas/camtrapr", build_vignettes = TRUE)
library(camtrapr)
```

# Use

This package is designed to take a set of identified and organized camera trap photos in a directory structure like this:

```
camera-trap-photos/
├── logged
│   ├── cam001
│   │   ├── muntjac
│   │   │   └── 1
│   │   │       ├── IMG020.JPG
│   │   │       ├── IMG021.JPG
│   │   │       └── IMG022.JPG
│   │   ├── squirrel
│   │   │   └── 1
│   │   │       ├── IMG025.JPG
│   │   └── wild-boar
│   │       ├── 1
│   │       │   ├── IMG032.JPG
│   │       └── 2
│   │           ├── IMG049.JPG
│   └── cam002
│       ├── squirrel
│       │   └── 1
│       │       ├── IMG034.JPG
│       └── wild-boar
│           ├── 2
│           │   ├── IMG046.JPG
│           └── x
│               ├── IMG048.JPG
└── primary
    ├── cam003
    │   ├── ignore
    │   │   └── IMG001.JPG
    │   ├── muntjac
    │   │   └── 1
    │   │       ├── IMG004.JPG
    │   └── porcupine
    │       └── 1
    │           └── IMG008.JPG
    └── cam004
        ├── muntjac
        │   └── 1
        │       ├── IMG010.JPG
        ├── squirrel
        │   └── 2
        │       ├── IMG012.JPG
        └── wild-boar
            └── 1
                ├── IMG015.JPG
```

And produce a tidy data frame ready for analysis like this:

```{r}
photo_path <- system.file("extdata", "example-photos", package = "camtrapr")
cam_data <- cam_process(photo_path, verbose = FALSE)
knitr::kable(head(cam_data, 5))
```

For further details on the use of this package consult the vignette:

```{r, eval=F}
browseVignettes('camtrapr')
```

